---
title: "LRS benchmarking"
author: "Jiangyan Yu (jiangyan.yu@ukw.de)"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# libraries

```{r,include=FALSE}
lapply(c("data.table","magrittr","ggplot2","dplyr","tidyr","vcfR","cowplot","ComplexHeatmap","UpSetR","scales","stringr","lubridate"), require, character.only = TRUE)


```

# working directory

```{r}
working_dir = "/Users/Yu_J/Documents/GitHub/lrs_benchmarking"


files = list.files(paste0(working_dir,"/truvari"))

files = files[files %like% "30x"]
files = files[!files %like% "log"]

```

# refine.regions.txt

## load all files

```{r}


## append F1 column from each file

df = lapply(files, FUN = function(f) {
  df_temp = read.table(paste0(working_dir,"/truvari/",f,"/refine.regions.txt"), header = TRUE)

  df_temp$seq_tech = str_split_fixed(f, "\\.", n = 6)[,2]
  df_temp$aligner = str_split_fixed(f, "\\.", n = 6)[,5]
  df_temp$SVcaller = str_split_fixed(f, "\\.", n = 6)[,6]
  df_temp$size = df_temp$end - df_temp$start + 1

  df_metrics <- df_temp %>%
    mutate(
      TP = in_tp,
      FP = in_fp,
      FN = in_fn,
      precision = ifelse((TP + FP) > 0, TP / (TP + FP), NA_real_),
      recall    = ifelse((TP + FN) > 0, TP / (TP + FN), NA_real_),
      F1        = ifelse((2*TP + FP + FN) > 0,
                         2*TP / (2*TP + FP + FN),
                         NA_real_)
    )

  ## append F1 to df_temp and rename the column name as f
  df_metrics$sample = f
  return(df_metrics)
}) %>% do.call(rbind,.)


```

## compare sv caller differences across regions

```{r}
## perform statisical test to compare F1 across different SV callers per region
df$region_id = paste(df$chr, df$start, df$end, sep = "_")
summary_state <- df %>%
  group_by(region_id) %>%
  summarise(
    chr = first(chrom),
    start = first(start),
    mean_F1 = mean(F1, na.rm = TRUE),
    median_F1 = median(F1, na.rm = TRUE),
    sd_F1 = sd(F1, na.rm = TRUE)
  )


```

### plot sd of F1 across regions

```{r, fig.width=10,fig.height=5}
## order the chr from chr1 to chr22
summary_state$chr <- factor(summary_state$chr, levels = paste0("chr", c(1:22)))


summary_state = summary_state[order(summary_state$start,decreasing = FALSE),]
summary_state = summary_state[order(summary_state$chr),]

## split by chromosome and plot sd of F1 across regions
p_sd <- ggplot(summary_state, aes(x = start, y = sd_F1,colour = chr)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "SD of F1 across regions")+
  facet_wrap(~chr, scales = "free_x",ncol=6)


p_sd
```

### histogram of sd

```{r, fig.width=5,fig.height=3}
ggplot(summary_state, aes(x = sd_F1)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  labs(title = "Distribution of SD of F1 across regions", x = "SD of F1", y = "Count")
```

## generally poor performance of SV callers across regions

```{r, fig.width=10,fig.height=5}
## plot the distribution of F1 across regions
ggplot(summary_state, aes(x = start, y = mean_F1,colour = chr)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Mean of F1 across regions")+
  facet_wrap(~chr, scales = "free_x",ncol=6)
```

## rm temp obj

```{r}
rm(df,summary_state,p_sd)
```

#---FN.vcf.gz and summary.json contains information before refine step---

# log.txt

```{r}
f1 = lapply(files, FUN= function(f){

  df = read.csv(paste0(working_dir,"/truvari/",f,"/log.txt"),header = FALSE)
  
  ## take the last 10 rows
  df = df[(nrow(df)-9):(nrow(df)-3),]
  ## remove space 
  df = df %>% mutate(V1 = str_replace_all(V1, " ", ""))
  df$parameter = str_split_fixed(df$V1, ":", n = 2)[,1]
  df$value = str_split_fixed(df$V1, ":", n = 2)[,2] %>% as.numeric()
  
  df1 = data.frame(setting = f,
                   precision = df[df$parameter == "precision", "value"],
                   recall = df[df$parameter == "recall", "value"],
                   F1 = df[df$parameter == "f1", "value"],
                   FP = df[df$parameter == "FP", "value"],
                   FN = df[df$parameter == "FN", "value"])
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]
  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  return(df1)
  
}) %>% do.call(rbind,.)
```


## F1

```{r,fig.height=3,fig.width=12}
ggplot(f1,aes(y=SVcaller,x=F1, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.8, 1)) +
  geom_vline(xintercept = 0.9, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "F1 score per setting", x = "F1 score")+
  facet_grid(seq_tech~aligner)
```

## F1

```{r,fig.height=8,fig.width=6}
df = f1
df = df[order(df$F1, decreasing = FALSE),]
df$setting = factor(df$setting, levels = df$setting)
ggplot(df,aes(y=setting,x=F1,fill=SVcaller))+
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() 
```

## precision

```{r,fig.height=3,fig.width=12}
ggplot(f1,aes(y=SVcaller,x=precision, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.9, 1)) +
  geom_vline(xintercept = 0.92, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "precision score per setting", x = "precision score")+
  facet_grid(seq_tech~aligner)
```

## recall

```{r,fig.height=3,fig.width=12}
ggplot(f1,aes(y=SVcaller,x=recall, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.7, 0.95)) +
  geom_vline(xintercept = 0.75, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "recall score per setting", x = "recall score")+
  facet_grid(seq_tech~aligner)
```

## FP


```{r,fig.height=3,fig.width=12}
ggplot(f1,aes(y=SVcaller,x=FP, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  # coord_cartesian(xlim = c(0.7, 0.95)) +
  # geom_vline(xintercept = 0.75, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "FP calls per setting", x = "Count")+
  facet_grid(seq_tech~aligner)
```


## FN


```{r,fig.height=3,fig.width=12}
ggplot(f1,aes(y=SVcaller,x=FN, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  # coord_cartesian(xlim = c(0.7, 0.95)) +
  # geom_vline(xintercept = 0.75, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "FN calls per setting", x = "Count")+
  facet_grid(seq_tech~aligner)
```


# FN.vcf.gz

```{r}


## append F1 column from each file

fn = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/fn.vcf.gz"))
  ### set the name of the list

  
  return(df_temp)
}) 

names(fn) = files
```

## SV length difference

TRF: tandom repeat finder

```{r}

fn_len = lapply(names(fn), FUN = function(f) {
  df1 = fn[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]
  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

fn_len_df = do.call(rbind,fn_len)


```


### total number of FN calls per setting

```{r,fig.height=3,fig.width=12}
df1 = table(fn_len_df$setting) %>% as.data.frame()
colnames(df1) = c("setting","count")

df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]


df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)

ggplot(df1,aes(y=SVcaller,x=count, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ## add a vertical line at 6000
  geom_vline(xintercept = 6000, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "Total number of FN calls per setting", x = "Sequencing technology and aligner", y = "Count")+
  facet_grid(seq_tech~aligner)
```


### histogram per setting

```{r,fig.height=8,fig.width=12}
ggplot(fn_len_df, aes(x = affected_length, fill = SVcaller)) +
   geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   scale_x_log10()+
  # scale_y_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of affected SV length in FN calls", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~SVcaller,scales = "free")

ggplot(fn_len_df,aes(x=affected_length, colour = SVcaller)) +
  geom_density() +
  theme_bw() +
  scale_x_log10()+
  theme(text = element_text(size = 12)) +
  labs(title = "Density of affected SV length in FN calls", x = "Affected SV length", y = "Density")+
  facet_grid(aligner~seq_tech, scales = "free_y")
```


### histogram per setting (<250)

```{r}



ggplot(subset(fn_len_df,affected_length<250),aes(x=affected_length, colour = SVcaller)) +
  geom_density() +
  theme_bw() +
  # scale_y_log10()+
  theme(text = element_text(size = 12)) +
  labs(title = "Density of affected SV length (<250) in FN calls", x = "Affected SV length", y = "Density")+
  facet_grid(aligner~seq_tech, scales = "free_y")
```

### histogram per setting (<1000)

```{r}

ggplot(subset(fn_len_df,affected_length<1000), aes(x = affected_length, fill = SVcaller)) +
   geom_bar(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of affected SV length (<1kb) in FN calls", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~SVcaller)

ggplot(subset(fn_len_df,affected_length<1000),aes(x=affected_length, colour = SVcaller)) +
  geom_density() +
  theme_bw() +
  # scale_y_log10()+
  theme(text = element_text(size = 12)) +
  labs(title = "Density of affected SV length (<1kb) in FN calls", x = "Affected SV length", y = "Density")+
  facet_grid(aligner~seq_tech, scales = "free_y")
```

## group SV length

```{r}
fn_len_df$length_group = cut(fn_len_df$affected_length, breaks = c(0, 50, 100, 250, 500, 1000,5000, 10000,20000,30000,Inf), 
                             labels = c("0-50", "50-100", "100-250", "250-500", "500-1000", "1k-5k","5k-10k","10k-20k","20k-30k",">30k"), right = FALSE)
```

### barplot of counts of each group

```{r, fig.width=10,fig.height=4}

  
ggplot(fn_len_df, aes(x = length_group, fill = SVcaller)) +
   geom_bar(position = "dodge") +
  scale_y_log10()+
   theme_bw() +
   theme(text = element_text(size = 12)) +
   labs(title = "Counts of FN calls in different length groups", x = "Affected SV length group", y = "Count")+
   facet_grid(aligner~seq_tech, scales = "free_y")
  

```

## any variants missed by all caller?

```{r}
fn_overlap = lapply(names(fn), FUN = function(f) {
  df1 = fn[[f]]@fix %>% as.data.frame()
  df1$var = paste0(df1$CHROM,"_",df1$POS,"_",df1$REF,"_",df1$ALT)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]
  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("var","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 


fn_overlap_df = do.call(rbind,fn_overlap)
```

### missed variants overlapped between SVcallers 

```{r}
## Upset plot to show the overlapped variants missed by different callers per seq_aligner
## per tech_aligner generate the upset figure
for(i in unique(fn_overlap_df$tech_aligner)) {
  df_temp = subset(fn_overlap_df, tech_aligner == i)
  
  listinput = list(sawfish=subset(df_temp,SVcaller=="sawfish")$var,
                   cuteSV=subset(df_temp,SVcaller=="cuteSV")$var,
                   pbsv=subset(df_temp,SVcaller=="pbsv")$var,
                   sniffles2=subset(df_temp,SVcaller=="sniffles2")$var)
  
  p = upset(fromList(listinput), 
            order.by = "freq", 
            mainbar.y.label = i,
            main.bar.color = "steelblue", 
            sets.bar.color = "steelblue", 
            text.scale = c(2, 2, 2, 1.5, 2, 1.5))
  print(p)

}


```

### pb-pbmm2

```{r}
df_temp = subset(fn_overlap_df, tech_aligner == "pb_pbmm2-pb")
```


# FP.vcf.gz

```{r}


fp = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/fp.vcf.gz"))

  
  return(df_temp)
}) 

names(fp) = files
```

## SV length difference

TRF: tandom repeat finder

```{r}

fp_len = lapply(names(fp), FUN = function(f) {
  df1 = fp[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

fp_len_df = do.call(rbind,fp_len)


```

### total number of FP calls per setting

```{r,fig.height=3,fig.width=12}
df1 = table(fp_len_df$setting) %>% as.data.frame()
colnames(df1) = c("setting","count")

df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]


df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)

ggplot(df1,aes(y=SVcaller,x=count, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  ## add a vertical line at 1500
  geom_vline(xintercept = 1500, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "Total number of FP calls per setting", x = "Sequencing technology and aligner", y = "Count")+
  facet_grid(seq_tech~aligner)
```


### histogram per setting

```{r,fig.height=8,fig.width=12}

ggplot(fp_len_df, aes(x = affected_length, fill = SVcaller)) +
   geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   scale_y_log10()+
  # scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of affected SV length in FP calls", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~SVcaller,scales = "free")

# ggplot(fp_len_df,aes(x=affected_length, colour = SVcaller)) +
#   geom_density() +
#   theme_bw() +
#   scale_x_log10()+
#   theme(text = element_text(size = 12)) +
#   labs(title = "Density of affected SV length in FP calls", x = "Affected SV length", y = "Density")+
#   facet_grid(aligner~seq_tech, scales = "free_y")
```

### cuteSV FP calls are mostly small SVs

```{r,fig.height=6,fig.width=5}

ggplot(fp_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of affected SV length in FP calls", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```

### ONT produces more FP large SVs

```{r,fig.height=6,fig.width=12}

ggplot(fp_len_df, aes(x = affected_length, color = seq_tech)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of affected SV length in FP calls", x = "Affected SV length group", y = "Density")+
   facet_grid(aligner~SVcaller,scales = "free")

```


## group SV length

```{r}
fp_len_df$length_group = cut(fp_len_df$affected_length, breaks = c(0, 50, 100, 250, 500, 1000, 5000,10000,20000,30000,Inf), 
                             labels = c("0-50", "50-100", "100-250", "250-500", "500-1000", "1k-5k","5k-10k","10k-20k","20k-30k",">30k"), right = FALSE)
```

### barplot of counts of each group

```{r, fig.width=12,fig.height=4}

  
ggplot(fp_len_df, aes(x = length_group, fill = SVcaller)) +
   geom_bar(position = "dodge") +
   theme_bw() +
  scale_y_log10()+
   theme(text = element_text(size = 12)) +
   labs(title = "Counts of FP calls in different length groups", x = "Affected SV length group", y = "Count")+
   facet_grid(aligner~seq_tech, scales = "free_y")
  
table(fp_len_df$setting,fp_len_df$length_group)
```

## rm temp obj

```{r}
## remove all objects
rm(list = ls())

```


# ---summary on files after refine step---

```{r}
working_dir = "/Users/Yu_J/Documents/GitHub/lrs_benchmarking"


files = list.files(paste0(working_dir,"/truvari"))

files = files[files %like% "30x"]
files = files[!files %like% "log"]
```

# refine.variant_summary.json

```{r}
## append F1 column from each file

refine_sum = lapply(files, FUN = function(f) {
  df1 = jsonlite::read_json(paste0(working_dir,"/truvari/",f,"/refine.variant_summary.json"))
  df1 = unlist(df1) %>% t() %>% as.data.frame()
  

  ## append F1 to df_temp and rename the column name as f
  df1$setting = f
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]
  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1)
  
}) %>% do.call(rbind,.)
```

## F1

```{r,fig.height=3,fig.width=12}
ggplot(refine_sum,aes(y=SVcaller,x=f1, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.8, 1)) +
  geom_vline(xintercept = 0.85, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "F1 score per setting (refine.variant_summary.json)", x = "F1 score")+
  facet_grid(seq_tech~aligner)
```

## precision

```{r,fig.height=3,fig.width=12}
ggplot(refine_sum,aes(y=SVcaller,x=precision, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.9, 1)) +
  geom_vline(xintercept = 0.92, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "precision score per setting (refine.variant_summary.json)", x = "precision score")+
  facet_grid(seq_tech~aligner)
```

## recall

```{r,fig.height=3,fig.width=12}
ggplot(refine_sum,aes(y=SVcaller,x=recall, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_cartesian(xlim = c(0.7, 0.95)) +
  geom_vline(xintercept = 0.75, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "recall score per setting (refine.variant_summary.json)", x = "recall score")+
  facet_grid(seq_tech~aligner)
```

## FP


```{r,fig.height=3,fig.width=12}
ggplot(refine_sum,aes(y=SVcaller,x=FP, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  # coord_cartesian(xlim = c(0.7, 0.95)) +
  geom_vline(xintercept = 1000, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "FP events per setting (refine.variant_summary.json)", x = "Count")+
  facet_grid(seq_tech~aligner)
```


## FN


```{r,fig.height=3,fig.width=12}
ggplot(refine_sum,aes(y=SVcaller,x=FN, fill = SVcaller)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  # coord_cartesian(xlim = c(0.7, 0.95)) +
  geom_vline(xintercept = 5000, linetype = "dashed", color = "red") +
  theme(text = element_text(size = 12),
        strip.text.y = element_text(angle = 0)) +
  labs(title = "FN events per setting (refine.variant_summary.json)", x = "Count")+
  facet_grid(seq_tech~aligner)
```

# refine.base.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

refine_base = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/refine.base.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(refine_base) = files
```

## length

```{r}
refine_base_len = lapply(names(refine_base), FUN = function(f) {
  df1 = refine_base[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

refine_base_len_df = do.call(rbind,refine_base_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(refine_base_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in refine.base.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```


# refine.comp.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

refine_comp = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/refine.comp.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(refine_comp) = files
```

## length

```{r}
refine_comp_len = lapply(names(refine_comp), FUN = function(f) {
  df1 = refine_comp[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

refine_comp_len_df = do.call(rbind,refine_comp_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(refine_comp_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in refine.comp.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```

# tp-base.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

tp_base = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/tp-base.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(tp_base) = files
```

## length

```{r}
tp_base_len = lapply(names(tp_base), FUN = function(f) {
  df1 = tp_base[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

tp_base_len_df = do.call(rbind,tp_base_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(tp_base_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in tp-base.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```


# tp-comp.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

tp_comp = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/tp-comp.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(tp_comp) = files
```

## length

```{r}
tp_comp_len = lapply(names(tp_comp), FUN = function(f) {
  df1 = tp_comp[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

tp_comp_len_df = do.call(rbind,tp_comp_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(tp_comp_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in tp-comp.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```

# fp.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

fp = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/fp.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(fp) = files
```

## length

```{r}
fp_len = lapply(names(fp), FUN = function(f) {
  df1 = fp[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

fp_len_df = do.call(rbind,fp_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(fp_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in fp.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```


# fn.vcf.gz

## load vcf files

```{r}
## append F1 column from each file

fn = lapply(files, FUN = function(f) {
  ## load vcf.gz file
  
  df_temp = read.vcfR(paste0(working_dir,"/truvari/",f,"/fn.vcf.gz"))
  
  ### set the name of the list

  
  return(df_temp)
}) 

names(fn) = files
```

## length

```{r}
fn_len = lapply(names(fn), FUN = function(f) {
  df1 = fn[[f]]@fix %>% as.data.frame()
  df1$ref_length = nchar(df1$REF)
  df1$alt_length = nchar(df1$ALT)
  df1$affected_length = apply(df1[,c("ref_length","alt_length")], 1, max)
  df1$setting = f
  
  
  df1$seq_tech = str_split_fixed(df1$setting, "\\.", n = 6)[,2]
  df1$aligner = str_split_fixed(df1$setting, "\\.", n = 6)[,5]
  df1$SVcaller = str_split_fixed(df1$setting, "\\.", n = 6)[,6]

  
  df1$tech_aligner = paste0(df1$seq_tech,"_",df1$aligner)
  
  
  return(df1[,c("affected_length","setting","seq_tech","aligner","SVcaller","tech_aligner")])
}) 

fn_len_df = do.call(rbind,fn_len)
```


## his

```{r,fig.height=6,fig.width=5}

ggplot(fn_len_df, aes(x = affected_length, color = SVcaller)) +
   # geom_histogram(position = "dodge") +
   ## overlay intensity plot using y axis on the right
   geom_density(aes(y = ..count..), alpha = 0.2, position = "identity") +
   theme_bw() +
   # scale_y_log10()+
  scale_x_log10()+
   theme(text = element_text(size = 12),
         strip.text.y = element_text(angle = 0)) +
   labs(title = "Density of SV length in fn.vcf.gz", x = "Affected SV length group", y = "Count")+
   facet_grid(tech_aligner~.,scales = "free")

```

# session info

```{r}
sessionInfo()
```

